Stack Safety as a Security Property

### 2021-03-30 tex/main.tex
Work on introduction and threat model

%% Abstract
%% Note: \begin{abstract}...\end{abstract} environment must come
%% before \maketitle command
\begin{abstract}
What is stack safety? Numerous mechanisms have been proposed to guard
against attacks that corrupt the call stacks of running programs,
\ifaftersubmission
and many
others rely on a protected stack as an assumption to provide higher level
guarantees,\bcp{this is a good point, but we don't pick it up anywhere in the
  paper, do we?}\leo{No, should we delete this? it still makes sense in flow}
\bcp{Let's delete for now; later, we can try to incorporate it in the body
  of the paper.}
\fi
but a clear and precise definition of stack safety is missing
from the literature.
%
We propose a formal definition phrased in the technical
terminology of language-based security---as a combination of a
confidentiality property (callees should not be able to read their caller's
stack frames) and an integrity property (callees should not corrupt their
caller's stack frames or control state), together with a straightforward
well-bracketedness condition on the program's control flow.
  %
We discuss how these properties are enforced or approximated by
existing stack protection mechanisms and describe useful
stronger and weaker variants.  \ifaftersubmission\bcp{should we mention stktokens here?}
\leo{Not sure. We almost never cite stuff in the abstract, do we?}\fi
\end{abstract}

@@

The call stack is a perennial target for attacks
based on memory-safety vulnerabilities, with potential consequences
ranging from leakage or corruption of private stack data to
control-flow hijacking. To detect or prevent such attacks, a menagerie of
software and hardware techniques for stack safety have been proposed,
%
including stack canaries~\citep{Cowan+98},
bounds checking~\citep{NagarakatteZMZ09,NagarakatteZMZ10,DeviettiBMZ08},
split stacks~\citep{Kuznetsov+14},
shadow stacks~\citep{Dang+15,Shanbhogue+19},
capabilities~\citep{Woodruff+14,Chisnall+15,Skorstengaard+19,Skorstengaard+19b},
and hardware tagging~\citep{DBLP:conf/sp/RoesslerD18}. \ifaftersubmission\apt{Mostly from
  nick; there could be more}\leo{Do we want to hunt for more refs here?}
\bcp{Sure, but later!}
\fi

But what {\em is} this stack safety of which people speak? Precise formal
definitions are mostly lacking: rather, the notion
is commonly defined in a negative way, by enumerating the bad things that a
given enforcement mechanism
prevents, such as corruption of return addresses, buffer overflows, or
use of uninitialized variables.
%\bcp{Does use of unitialized variables lead to stack safety attacks?\apt{sure, via buffer overflow}}
%with the notion usually being
%defined negatively: through examples of stack-based exploits
%\rb{usually though not always; as noted below there are other
%  approaches}\bcp{Yes, we have to be very careful about this}.
%\leo{Examples at a high level}: such as -- does not overwrite return addresses, or ...
Indeed, we are only aware of one quite recent attempt to frame a
positive definition, by
\citet{Skorstengaard+19}, which defines stack safety as
the conjunction of two properties: {\em local state encapsulation} and
{\em well-bracketed control flow}.
These properties are formalized by defining a capability-based machine
that ``satisfies [stack safety] by construction.''
%
While this work represents an important step in the right direction, we believe
there is still room to craft an operational characterization
of stack safety that is independent of capabilities or any other specific
enforcement technique.  In other words, we seek formal criteria that can be applied
to a mechanism for protecting stack-based low-level programs to judge
whether or not it actually ``guarantees stack safety.''
%
Our key contribution is an abstract presentation of local state encapsulation
using the technical framework of language-based security, together with a simple
characterization of well-bracketed control flow.

@@

When does the programmer rely on stack safety? The most natural example is the first
in figure \ref{fig:examples1}, in which {\tt main} calls stores 42 to a local variable {\tt x},
calls {\tt f}, and then prints {\tt x}. Here the programmer should be able to expect that
{\tt main} will print 42, but in C such an assumption may be wrong, if {\tt f} should
overwrite it's caller's stack data. To this end, our properties support a
{\em call rule} that can be used to reason that on return from {\tt f}, {\tt x} still
contains the same value. This is not only useful for the programmer: consider the second
example, where the call to {\tt h} should be unreachable and therefore subject to dead code
elimination, if only the compiler can soundly reason that {\tt x} does not change.
The property that enables the call rule is, in security terms, {\em integrity}: callers'
data is protected from modification by callees.

Alongside integrity, we have {\em confidentiality}, in which callers' data is protected
from leakage by their callees. Consider again the first example in figure \ref{fig:examples1},
in which {\tt f}'s behavior should not depend on {\tt x} being 42. In fact suppose that the
programmer had initialized {\tt x} after calling {\tt f} rather than before; such a change
should not alter the behavior of {\tt main}, but if {\tt f} can directly read {\tt x}, it
may. But there are more subtle examples, such as the first in figure \ref{fig:examples2},
where {\tt main} may call {\tt g} or {\tt h} are called prior to calling {\tt f}, depending on
its argument. Here, if {\tt g} or {\tt h} leaves behind identifying information in unused stack
space, {\tt f} could still deduce information about the argument and behave differently, violating
the abstraction of its call. So our confidentiality properties also support a {\em insensitivity rule},
that each function's behavior is independent of the state of its caller, other than arguments.
This also makes it easier, though not trivial, to prove a function's idempotence: the channels through
which a callee can remember its own history are limited to those outside the stack.

Together, integrity and confidentiality comprise \citeauthor{Skorstengaard+19}'s
{\em local state encapsulation} property. But LSE alone is not sufficient to fully protect
callers from their callees. Consider the last example in figure \ref{fig:examples2},
a variant of the famous ``awkward example,'' in which {\tt f} is called twice and may not
be idempotent. It should be the case that {\tt main} prints -1, but if
{\tt f}'s first activation stores its return address somewhere accessible, the second
activation can return to before {\tt x} is negated, causing {\tt x} to be set to
1 again before a third activation that would return normally. Then {\tt main} would print 1.
Once again this is rather unexpected behavior.

Preventing this form of reentrancy attack is the intuition behind
\citeauthor{Skorstengaard+19}'s {\em well-bracketed control flow}, which protects
the call graph itself. A callee should pass control back to the caller only by
transferring control to the instruction following the call; it should not be allowed
to jump back into the caller's code at an arbitrary point, either directly (via a jump
instruction) or indirectly (by corrupting stack data and returning).

To phrase LSE and WBCF as security properties requires an explicit attacker model:
a clear understanding of who is being protected and from what. Most existing
stack protection mechanisms envisage scenarios where a callee is
influenced by external inputs to corrupt the stack in some way. To
avoid getting into the details of exactly how this happens when
formalizing our properties, we use an even stronger model: when we
state the protections that a function can rely on when it makes a
call, we assume that the callee can execute arbitrary instructions
until it returns.

A key challenge in adapting the concepts of language-based security to this
setting is that integrity and confidentiality are ``nested'' properties: not
only does a given caller need to be protected from its callee, but the callee
also needs to be protected from {\em its} callees, and so on. That is, the notion
of ``attacker'' changes at many points throughout the execution of the program.
We address this issue by introducing a dynamically model of {\em domains} for
the different components of the machine state, in which some stack locations are
sealed by the caller that needs to preserve them while the rest are unsealed and
available for use.

First we introduce this model for a single simple stack with no sharing between callers and
callee. Then we expand it to allow components to be passed, granting a callee one time
access to the caller's data, or shared, allowing components to be outright
exempted from the stack safety properties. Shared components could model objects whose
addresses are passed and whose security must be modeled by a separate property, such as
memory safety enforced by capabilities.

After formally defining stack safety, we show that the definition is
enforceable using an existing tag-based enforcement
mechanism~\citep{DBLP:conf/sp/RoesslerD18}. We first consider
\citeauthor{DBLP:conf/sp/RoesslerD18}'s conservative Depth Isolation policy. 
We give {\em eager} properties that this policy enforces, in which each
callee, at the point of its return, can be seen to respect the integrity
and confidentiality of its caller. Additionally we have an
{\em ultra-eager} integrity property which guarantees that integrity
is respected at each step. The advantage of this property is that it ``fails
faster'' and is therefore more suitable for efficient testing. We prove (in
Coq) that the ultra-eager property implies our original definition of stack
safety.

The downside of the Depth Isolation policy is that it is rather
expensive to implement. In the same work,
\citeauthor{DBLP:conf/sp/RoesslerD18} propose a different mechanism
that is more efficient by virtue of being lazy in its enforcement.
Instead of enforcing the integrity and confidentiality of the stack at the
exact moment when a callee returns, this policy signals violations when the
caller actually accesses data that the callee has written into its stack
frame\ifaftersubmission\bcp{We could also hint at the lazy confidentiality
  property in the same way...}\fi.
The lazy policy does not enforce our original integrity and confidentiality
properties, and to our knowledge there is no formal statement of the properties that
it ought to enforce. We propose a {\em lazy} integrity property, which capture the
desired behavior of a policy that detects violations between when they occur and when
they make a visible impact. Unfortunately \citeauthor{DBLP:conf/sp/RoesslerD18}'s
existing ``Lazy Tagging and Clearing'' policy does not quite enforce these properties,
but formalizing the ideal property allows us to see how it falls short.
We provide a variant of ``Lazy Tagging and Clearing'' that does enforce lazy integrity
at some performance cost.

We extend our properties to allow sharing, both in the form of arguments passed
on the stack and of objects shared widely beyond their owning function. In the latter
case, restricting access the shared objects is beyond the scope of stack safety,
but our model delimits them from other objects in the same frame.

@@

In summary, we offer the following contributions:
\begin{itemize}
\item
  % \leo{More? Extensional}\bcp{We might need to invent a word for this---''quasi-extensional''?}
  A novel {\em stack-safety property} abstractly capturing the
  intuitive concepts of strict local state encapsulation and well-bracketed
  control flow~(\cref{sec:lse-and-wbcf}).
  % These properties rule out common examples of stack memory abuse.
  This property can be enforced by
  the Depth Isolation policy of \citet{DBLP:conf/sp/RoesslerD18}, which is
  based on hardware tagging~(\cref{sec:conservative}).
\item
  A {\em lockstep} variant of the stack-safety property that is also
  enforced by the Depth Isolation policy
  but is more efficient to test, plus a proof (in Coq) that this variant is
  stronger than the original~(\cref{sec:lockstep}).
\item
  A {\em lazy} variant of the stack-safety property that can be enforced if the
  Lazy Tagging and Clearing policy of \citet{DBLP:conf/sp/RoesslerD18} is enhanced
  with unique tags on each activation, plus a proof (in Coq) that this
  variant is weaker than the original~(\cref{sec:lazy}).
%  This policy does not enforce our strict
%  local state encapsulation property,
%  but a modified version of it using unique per-activation tags enforces
%  an (also provably) weaker lazy variant of the property, which still prohibits
%  the same intuitively stack unsafe behaviors.
%
% \apt{Awkward, but the truth I fear.} \bcp{I agree the phrasing is awkward,
%   but the point we can make is still strong: To make the property easier
%   to enforce, we can slightly weaken its guarantees; the weaker version is
%   a bit more complicated to state but still retains the pragmatic
%%   protections of the original.}\apt{To clarify: Nick's LT\&C policy uses
%   depth-based tags, which do not enforce even our lazy policy. So we need to
%   introduce a version of LT\&C that uses unique activation tags.}

@@

Stack safety is meant to protect a caller from a confused or malicious callee's
attempts to access private data or hijack its control flow. Our
attacker model reflects this intention by assuming that the attacker
can run arbitrary machine code to try to access or overwrite its caller's stack data
or disrupt its control state.  Hardware and timing attacks are out of
scope.
%\leo{Anything else to add here?}
%\leo{Maybe a good place to discuss calling conventions and that they
%  are essentially enforced by the policy or sth?}
%\rb{Intuition that the callee can do anything in principle, but the machine will
%halt if it does not follow the calling convention, etc.?}

@@

We assume we are given a machine language program, together with some auxiliary information
about its structure: the code footprint of each function and the locations of instructions
representing calls, returns, and function entry points.
Although this information would typically come from a compiler together with
the machine code, we do not
assume anything about their provenance.  We assume that the machine has a
program counter $\PCname$
and stack pointer $\rsp$; other details of the machine's ISA are unimportant.

To make our stack safety definitions precise, we make some straightforward
assumptions about
memory layout; our definitions could easily be changed to be parametric over
these assumptions, at the cost of some complexity in their statements.  Memory
consists of an instruction region and a stack that grows upwards. (For simplicity,
we do not consider the heap. Integrating a memory-safety property for the
heap in the style of
\citet{DBLP:conf/post/AmorimHP18} is a natural extension that we discuss
in~\Cref{sec:relwork}.)  The stack has a conventional structure, with
contiguously allocated per-function frames.
The boundary between the frames of a caller and callee is indicated by the $\rsp$ register
at the time of the call.  The stack can be used to
pass arguments or return results in the usual way, by placing them
at the top of the caller's frame. Callers and callees can also communicate
via registers if they wish; indeed, our
definitions say nothing at all about registers (except $\PCname$ and
$\rsp$). \ifaftersubmission\bcp{This implicitly amounts to choosing a caller-saves register
  convention, though, doesn't it?\apt{I think it is neutral; we could add separate caller register protection if we wanted.}}\fi

Our stack-safety concepts depend critically on the notions of ``call'' and ``return,''
which are not fully explicit in machine code (the same instruction opcode may be
used in some places as part of a call or return sequence and in other places
for other purposes). We assume the compiler's cooperation in labeling instructions
with a variety of annotations, and in fact in identifying them as instructions.
A {\em call-map} is a partial map from addresses to pairs of a function identifier --
every instruction belongs to a function -- and a set of annotations. This gives
us flexibility to have instructions simultaneously play multiple roles.

Enforcement mechanisms are modeled as {\em policies} (borrowing the
terminology from tag-based mechanisms) that add some auxiliary policy
state to the ordinary machine state
%\bcp{This may be surprising to people who are not thinking about micropolicies.\apt{added para below}}
and a partial policy step function that refines the machine step
function by allowing or disallowing a step to be taken based on its
auxiliary state.  Enforcement works by halting the program before it
can perform an action that would violate the security property; our
properties are therefore naturally \emph{termination insensitive.}
This model will be made more precise in \cref{sec:prelim}.

We model ordinary program state and policy-enforcement state separately to make it
easier to use well-established concepts from language-based security.
In particular, confidentiality has an elegant extensional characterization in terms of
\emph{noninterference}: a program preserves the confidentiality of certain data if
\emph{varying} that data would not change the program's observable
behavior~\citep{6234468}.
In applying this idea to a system that incorporates policy enforcement, only
ordinary program data should be varied, while any associated policy
state should be left alone; our model makes this distinction obvious.
%That is, we compare the actual execution of the program with a \emph{hypothetical}
%execution where data has been varied.
We will also introduce a notion of {\em observational integrity} that relies on the idea of
\emph{rolling back} program state; again, our model makes it easy to describe
rollback of ordinary program data while leaving policy state alone.

@@

This enforcement model directly describes
micropolicies~\citep{pump_oakland2015}, and we believe that it is
flexible enough to capture code-altering approaches (such as code rewriting
or capability-based techniques).

@@

\paragraph*{Software-only policies}
%
Some policies are implemented in software, by modifying code in the
compiler, for example to perform bounds checking~\citep{NagarakatteZMZ09} or
to insert stack canaries~\citep{Cowan+98}. Such approaches can, in principle,
be evaluated in terms of our formal notion of stack safety with a trivial null
policy. But any state maintained by the enforcement mechanism must be in normal
memory and subject to the same confidentiality and integrity concerns (to be
varied, rolled back, etc.), a heavy constraint. Alternately it can be stripped
out as policy state provided that it can be proven to protect itself.

@@

\subsection{Eager Integrity and the Call Rule}

\definition Let \(\mpcstatename\) be a reachable state whose PC value is annotated
as a call and whose call depth is \(\depth\). Assume \(\mpcstatename \stepstocon \mpcstatename'\).
Let \(\MPCT\) be the trace ending after the corresponding return, if any, defined by
\(\mpcstatename' \hookrightarrow \MPCT | f\) with \(f ~ \mach ~ \context\) holding when
\(\context = \_,\depth\). Finally let \(\components\) be the set of components \(\component\)
such that \(\sealcon ~ \pi_\mach(\mpcstatename) ~ \component\) holds. Then by the {\em Call Rule},
it is sound to assume that if \(\last(\pi_\mach(\MPCT)) = \mach'\), then
\(\pi_\mach(\mpcstatename) ~ \component = \mach' ~ \component\).

@@

\subsection{Eager Confidentiality and the Insensitivity Rule}

@@

\definition Let \(\mpcstatename\) be a reachable call state. Assume the that the call is successful
and \(\mpcstatename \stepstocon \mpcstatename'\). Then let \(\npcstatename'\) be a \(\components\)-variant
where \(\components\) is the set of all addresses in the stack. Then by the {\em Insensitivity Rule},
if the trace from \(\mpcstatename'\) eventually returns, so does that from \(\npcstatename'\),
their states after the return have the same differences, and up to return their traces are observationally
equivalent.

@@

\section{Programming Logic Perspective}

@@

\section{Stack Safety Properties for Lazy Enforcement}
\label{sec:lazy}

@@

  \paragraph{Stack-derived Pointers}

    Common programming idioms involve a caller passing a pointer to
    its local data into a callee. Using standard notions of pointer
    provenance~\citep{provenance}, we can extend our model to
    distinguish safe use of this idiom from true stack safety
    violations. We require additional annotation to identify which
    addresses within a function's stack frame correspond to distinct
    objects. Then a valid pointer to an object is one derived through
    legal arithmetic from its base address, and at each call point, an
    object's addresses are marked low confidentiality and integrity if
    a valid pointer exists in a register or in the transitive closure
    of the callee's accessible memory. All pointers to an object cease
    to be valid when it is deallocated along with its stack frame.

  \paragraph{Non-stack control flow}

    Control structures beyond well-bracketed calls and returns require
    modifications to the model. Tail calls, for instance, reuse the caller's
    stack frame for its callee, and every nested tail call returns
    simultaneously to the top non-tail-calling function. Under normal stack
    integrity, the first function in a chain of tail calls returns and violates
    integrity, while subsequent tail calls never actually return at all.
    Instead we need a variant that treats the entry of a tail call as the
    return of its caller.

>>>>

%% Abstract
%% Note: \begin{abstract}...\end{abstract} environment must come
%% before \maketitle command
\begin{abstract}
  What is stack safety? Numerous mechanisms have been proposed to guard against
  attacks that corrupt the call stacks of running programs, including micro-policies
  implemented using metadata tagging mechanisms such as PIPE. But for specification
  these implementations largely appeal either to intuitive notions of the behavior
  of a stack, or to specific examples of bad behavior.
  %
  We propose a formal definition phrased in the technical terminology of language-based
  security---as a combination of an integrity property (a caller's stack frame is
  invariant over calls) and a confidentiality property (callees lack knowledge of the
  contexts of their calls). These combine with a straightforward well-bracketedness
  condition on the program's control flow that ensures that callers and callees remain
  meaningfully distinct.
  %
  While our formalism is informed by the PIPE implementation, the primary properties reflect
  general principles of stack safety. We define additional properties tailored for specific
  contexts: a stronger property amenable to efficient testing and a weaker one that is
  approximated by more efficient, lazy policies.

  \ifaftersubmission\bcp{should we mention stktokens here?}
  \leo{Not sure. We almost never cite stuff in the abstract, do we?}
  \sna{On the same note, cite Nick and Andre?}
  \fi
\end{abstract}

@@

The call stack is a perennial target for attacks
based on memory-safety vulnerabilities, with potential consequences
ranging from leakage or corruption of private stack data to
control-flow hijacking. To detect or prevent such attacks, a menagerie of
software and hardware techniques for stack safety have been proposed,
%
including stack canaries~\citep{Cowan+98},
bounds checking~\citep{NagarakatteZMZ09,NagarakatteZMZ10,DeviettiBMZ08},
split stacks~\citep{Kuznetsov+14},
shadow stacks~\citep{Dang+15,Shanbhogue+19},
capabilities~\citep{Woodruff+14,Chisnall+15,Skorstengaard+19,Skorstengaard+19b},
and hardware tagging~\citep{DBLP:conf/sp/RoesslerD18}. \ifaftersubmission\apt{Mostly from
  nick; there could be more}\leo{Do we want to hunt for more refs here?}
\bcp{Sure, but later!}
\fi
This last provides an array of micro-policies offering protection at various granularities,
which offer a trade-off between protection and efficiently.

But what {\em is} this stack safety of which they speak? Precise formal
definitions are mostly lacking: rather, the notion is commonly defined in a
negative way, by enumerating the bad things that a given enforcement mechanism
prevents, such as corruption of return addresses, buffer overflows, or
use of uninitialized variables.
%\bcp{Does use of unitialized variables lead to stack safety attacks?\apt{sure, via buffer overflow}}
%with the notion usually being
%defined negatively: through examples of stack-based exploits
%\rb{usually though not always; as noted below there are other
%  approaches}\bcp{Yes, we have to be very careful about this}.
%\leo{Examples at a high level}: such as -- does not overwrite return addresses, or ...
Indeed, we are only aware of one quite recent attempt to frame a positive definition, by
\citet{Skorstengaard+19}, which defines stack safety as the conjunction of two properties:
{\em local state encapsulation} and {\em well-bracketed control flow}.
These properties are formalized by defining a capability-based machine
that ``satisfies [stack safety] by construction.''
%
While this work represents an important step in the right direction, we believe
there is still room to craft an operational characterization
of stack safety. In other words, we seek formal criteria that can be applied
to a mechanism for protecting stack-based low-level programs to judge
whether or not it actually ``guarantees stack safety.'' While our criteria are
designed with PIPE micro-policies in mind, they are agnostic about many implementation
details and should be applicable to other systems with minor adjustment.
%
Our key contribution is an abstract presentation of local state encapsulation
using the technical framework of language-based security, together with a simple
characterization of well-bracketed control flow.

@@

When does the programmer rely on stack safety? The most natural example is the first
in figure \ref{fig:examples1}, in which {\tt main} calls stores 42 to a local variable {\tt x},
calls {\tt f}, and then prints {\tt x}. Here the programmer should be able to expect that
{\tt main} will print 42, but in C such an assumption may be wrong, if {\tt f} should
overwrite {\tt main}'s stack data. This expectation is for {\em integrity}: that a
caller's data is protected from modification by its callee.
This is not only useful for the programmer: consider the second
example, where the call to {\tt h} should be unreachable and therefore subject to dead code
elimination, if only the compiler can soundly reason that {\tt x} does not change.

Alongside integrity, we have {\em confidentiality}, in which a caller's data is protected
from leakage by its callee. Consider again the first example in figure \ref{fig:examples1},
in which {\tt f}'s behavior should not depend on {\tt x} being 42. In fact suppose that the
programmer had initialized {\tt x} after calling {\tt f} rather than before; such a change
should not alter the behavior of {\tt main}, but if {\tt f} can directly read {\tt x}, it
may. But there are more subtle examples, such as the first in figure \ref{fig:examples2},
where {\tt main} may call {\tt g} or {\tt h} are called prior to calling {\tt f}, depending on
its argument. Here, if {\tt g} or {\tt h} leaves behind identifying information in unused stack
space, {\tt f} could still deduce information about the argument and behave differently, violating
the abstraction of its call.

Together, integrity and confidentiality comprise \citeauthor{Skorstengaard+19}'s
{\em local state encapsulation} property. But LSE alone is not sufficient to fully protect
callers from their callees. Consider the last example in figure \ref{fig:examples2},
a variant of the famous ``awkward example,'' in which {\tt f} is called twice and might store
data outside the stack between calls. It should be the case that {\tt main} prints -1, but if
{\tt f}'s first activation stores its return address somewhere accessible, the second
activation can return to before {\tt x} is negated, causing {\tt x} to be set to
1 again before a third activation that would return normally. Then {\tt main} would print 1.
This is rather unexpected behavior.

Preventing this form of control-flow attack is the intuition behind
\citeauthor{Skorstengaard+19}'s {\em well-bracketed control flow}, which protects
the call graph itself. A callee should pass control back to the caller only by
transferring control to the instruction following the call; otherwise it could
subvert LSE by appearing to return -- thus restoring access to the caller's data --
while causing the execution of arbitrary instructions on that data.

To phrase LSE and WBCF as security properties requires an explicit attacker model:
a clear understanding of who is being protected and from what. Most existing
stack protection mechanisms envisage scenarios where a callee is
influenced by external inputs to corrupt the stack in some way. To
avoid getting into the details of exactly how this happens when
formalizing our properties, we use an even stronger model: when we
state the protections that a function can rely on when it makes a
call, we assume that the callee can execute arbitrary instructions
until it returns. Thus we have no need to differentiate safe code
from adversarial - all functions, and indeed all activations of the same function,
are mutually distrusting.
%
Even more fundamentally, we need to know when a call or a return is taking
place. We assume that calls and returns are annotated in the code, so that
we can track the point at which control transfers from the protected caller
to the attacking callee -- and when that callee in turn becomes a protected
caller in its own right.

This ``nesting'' of properties is a key challenge in adapting the concepts
of language-based security to the stack setting. We address this issue by
turning any set of annotations into a dynamic model of {\em domains},
in which some stack locations are sealed by the caller that needs to preserve
them while the rest are unsealed and available for use by the active, attacking
callee.

First we introduce this model for a single simple stack with no sharing between callers and
callee. Then we expand it to allow components to be passed, granting a callee one time
access to the caller's data, or shared, allowing components to be outright
exempted from the stack safety properties. Shared components could model objects whose
addresses are passed and whose security must be modeled by a separate property, such as
memory safety enforced by capabilities. Finally, we extend the model to a coroutine
system featuring finitely-bounded stacks.

After formally defining stack safety, we show that the definition is
enforceable using an existing tag-based enforcement
mechanism~\citep{DBLP:conf/sp/RoesslerD18}. We first consider
\citeauthor{DBLP:conf/sp/RoesslerD18}'s conservative Depth Isolation policy. 
This policy testably enforces our {\em eager} properties, in which each
callee, at the point of its return, can be seen to respect the integrity
and confidentiality of its caller. Additionally we have an
{\em ultra-eager} integrity property which guarantees that integrity
is respected at each step. The advantage of this property is that it ``fails
faster'' and is therefore more suitable for efficient testing. We prove (in
Coq) that the ultra-eager property implies our original definition of integrity.

The downside of the Depth Isolation policy is that it is rather
expensive to implement. In the same work,
\citeauthor{DBLP:conf/sp/RoesslerD18} propose a different mechanism
that is more efficient by virtue of being lazy in its enforcement.
Instead of enforcing the integrity and confidentiality of the stack at the
exact moment when a callee returns, this policy signals violations when the
caller actually accesses data that the callee has written into its stack
frame. \ifaftersubmission\bcp{We could also hint at the lazy confidentiality
  property in the same way...}\fi
The lazy policy does not enforce our original integrity property, and to our
knowledge there is no formal statement of the properties that it ought to enforce.
We propose a {\em lazy} integrity property, which capture the desired behavior of a
policy that detects integrity violations between when they occur and when they make
a visible impact. Intuitively this can be phrased, ``No visible reads after dangerous
writes.'' Unfortunately \citeauthor{DBLP:conf/sp/RoesslerD18}'s existing
``Lazy Tagging and Clearing'' policy does not quite enforce this property, as shown in
testing, but formalizing the ideal property allows us to see how it falls short.
We provide a variant of ``Lazy Tagging and Clearing'' that does enforce lazy integrity
at some performance cost.

@@

In summary, we offer the following contributions:
\begin{itemize}
\item
  % \leo{More? Extensional}\bcp{We might need to invent a word for this---''quasi-extensional''?}
  A novel {\em stack-safety property} abstractly capturing the
  intuitive concepts of strict local state encapsulation and well-bracketed
  control flow~(\cref{sec:lse-and-wbcf}).
  % These properties rule out common examples of stack memory abuse.
\item
  A property-based randomized testing framework that demonstrates that this
  property can be enforced by the Depth Isolation policy of
  \citet{DBLP:conf/sp/RoesslerD18}, which is based on hardware tagging~(\cref{sec:conservative}).
\item
  An {\em ultra-eager} variant of the stack-integrity property that is also
  enforced by the Depth Isolation policy but is more efficient to test, plus a proof
  (in Coq) that this variant is stronger than the original~(\cref{sec:lockstep}).
\item
  A {\em lazy} variant of the stack-safety property that can be enforced if the
  Lazy Tagging and Clearing policy of \citet{DBLP:conf/sp/RoesslerD18} is enhanced
  with unique tags on each activation, plus a proof (in Coq) that this
  variant is weaker than the original~(\cref{sec:lazy}).
\item
  Standardized trace-based formulations of eager confidentiality and of ultra-eager,
  eager, and lazy integrity, parameterizeable by arbitrary control-flow models and assignments
  of machine components to {\em LSE domains}.
\item
  An LSE-domain model tracking the passing of function arguments on the stack and the sharing
  of arbitrary stack data.
\item
  An extension of stack safety to a simple coroutine model with corresponding LSE
  properties and a well-structured control flow property that subsumes WBCF for each
  stack as well as guaranteeing legitimate coroutine crossings.
%  This policy does not enforce our strict
%  local state encapsulation property,
%  but a modified version of it using unique per-activation tags enforces
%  an (also provably) weaker lazy variant of the property, which still prohibits
%  the same intuitively stack unsafe behaviors.
%
% \apt{Awkward, but the truth I fear.} \bcp{I agree the phrasing is awkward,
%   but the point we can make is still strong: To make the property easier
%   to enforce, we can slightly weaken its guarantees; the weaker version is
%   a bit more complicated to state but still retains the pragmatic
%%   protections of the original.}\apt{To clarify: Nick's LT\&C policy uses
%   depth-based tags, which do not enforce even our lazy policy. So we need to
%   introduce a version of LT\&C that uses unique activation tags.}

@@

Stack safety is meant to protect a caller from a confused or malicious callee's
attempts to access private data or hijack its control flow. And in fact the caller,
while presumably not malicious, may itself be confused or erroneous. Our attacker
model reflects this intention by assuming that both the attacker and the defender
may run arbitrary machine code. Hardware and timing attacks are out of scope.

%\leo{Anything else to add here?}
%\leo{Maybe a good place to discuss calling conventions and that they
%  are essentially enforced by the policy or sth?}
%\rb{Intuition that the callee can do anything in principle, but the machine will
%halt if it does not follow the calling convention, etc.?}

@@

In this very strong attacker model, there is little obvious distinction between
the attacker and the defender. They are separated only by the control transfer
inherent in their calls and returns. Our stack-safety concepts therefore depend
critically on the notions of ``call'' and ``return,'' which are not fully explicit
in machine code (the same instruction opcode may be used in some places as part of
a call or return sequence and in other places for other purposes).
%
We assume we are given a machine language program annotated with the locations of
instructions representing calls, returns, and function entry points. Although this
information would typically come from a compiler together with the machine code,
we do not assume anything about their provenance. We do treat the execution of an
annotated instruction as in some sense intentional on the part of the current
function. No accidental calls or returns, and attackers control their own returns
but cannot change the annotations on the caller's code.

\paragraph{Well-formedness and Bad Annotations}

Since we rely so heavily on code annotations, we must ask: what if the annotations
are wrong? What if a call or return occurs that is not annotated, or an annotation
fails to correspond to an actual return? In the former case, the answer is simple --
there has been no call or return, only instruction resembling such an operation, and
therefore no associated change in permissions. This is particularly interesting in the
case of a false return, where control may appear to have returned to the caller, but
the enforcement mechanism must still protect the caller's memory.

The latter case depends on the enforcement mechanism and the nature of the false
annotation. A given enforcement mechanism might place a static well-formedness criterion
on the code it applies to. For instance, our micro-policies expect that call and return
annotations occur on jump instructions. On the other hand, not all jumps are proper
calls or returns, only those preceded by appropriate sequences: in these cases the
bad annotation might appear well formed but the system should either respect the change
in permissions or, if doing so would compromise security, fail-stop rather than
execute the annotated instruction.

\paragraph{System Assumptions}

Our primary target is a RISC-V architecture extended with the PIPE architecture, but the
model makes very few assumptions about the machine it applies to. We assume a register
machine with a program counter \(\PCname\) and stack pointer \(\rsp\); other details of
the machine's ISA are unimportant.

Our definitions are largely parametric over the layout of the stack in memory, but for
purposes of explanation we will describe a standard layout. Memory consists of an
instruction region and a stack that grows upwards. (For simplicity,
we do not consider the heap. Integrating a memory-safety property for the
heap in the style of \citet{DBLP:conf/post/AmorimHP18} is a natural extension
that we discuss in~\Cref{sec:relwork}.) The stack has a conventional structure, with
contiguously allocated per-function frames.

The boundary between the frames of a caller and callee is indicated by the $\rsp$ register
at the time of the call. The stack can be used to pass arguments or return results
in the usual way, by placing them at the top of the caller's frame. This will be modeled
as a form of sharing in section \sna{cite the sharing section}.
Callers and callees can also communicate via registers if they wish; indeed, our
definitions say nothing at all about registers (except $\PCname$ and $\rsp$).

\ifaftersubmission\bcp{This implicitly amounts to choosing a caller-saves register
  convention, though, doesn't it?\apt{I think it is neutral; we could add separate caller register protection if we wanted.}}\fi

\paragraph{Policies}

We model the base machine as a step function over machine states, and a machine enhanced with
an enforcement mechanism (in the language of tag-based enforcement, a {\em policy}) as 
a partial step function over machine states extended with some auxiliary state, refining
the base machine. When the policy-enhanced machine fails to step, this is a {\em failstop},
halting the program before it can perform an action that would violate the security property; our
properties are therefore naturally \emph{termination insensitive.}
This model will be made more precise in \cref{sec:prelim}.

%We model ordinary program state and policy-enforcement state separately to make it
%easier to use well-established concepts from language-based security.
%In particular, confidentiality has an elegant extensional characterization in terms of
%\emph{noninterference}: a program preserves the confidentiality of certain data if
%\emph{varying} that data would not change the program's observable
%behavior~\citep{6234468}.
%In applying this idea to a system that incorporates policy enforcement, only
%ordinary program data should be varied, while any associated policy
%state should be left alone; our model makes this distinction obvious.
%That is, we compare the actual execution of the program with a \emph{hypothetical}
%execution where data has been varied.
%We will also introduce a notion of {\em observational integrity} that relies on the idea of
%\emph{rolling back} program state; again, our model makes it easy to describe
%rollback of ordinary program data while leaving policy state alone.

@@

This enforcement model directly describes micropolicies~\citep{pump_oakland2015}.
We believe that it is flexible enough to capture code-altering approaches, such as code
rewriting or capability-based techniques, perhaps with less elegance.

@@

@@

\subsection{Eager Integrity}

@@

\subsection{Eager Confidentiality}

@@

@@

@@

\section{Stack Safety for Lazy Enforcement}
\label{sec:lazy}

@@

  %\paragraph*{Software-only policies}
  %
  %Some policies are implemented in software, by modifying code in the
  %compiler, for example to perform bounds checking~\citep{NagarakatteZMZ09} or
  %to insert stack canaries~\citep{Cowan+98}. Such approaches can, in principle,
  %be evaluated in terms of our formal notion of stack safety with a trivial null
  %policy. But any state maintained by the enforcement mechanism must be in normal
  %memory and subject to the same confidentiality and integrity concerns (to be
  %varied, rolled back, etc.), a heavy constraint. Alternately it can be stripped
  %out as policy state provided that it can be proven to protect itself.

%  \paragraph{Stack-derived Pointers}

%    Common programming idioms involve a caller passing a pointer to
%    its local data into a callee. Using standard notions of pointer
%    provenance~\citep{provenance}, we can extend our model to
%    distinguish safe use of this idiom from true stack safety
%    violations. We require additional annotation to identify which
%    addresses within a function's stack frame correspond to distinct
%    objects. Then a valid pointer to an object is one derived through
%    legal arithmetic from its base address, and at each call point, an
%    object's addresses are marked low confidentiality and integrity if
%    a valid pointer exists in a register or in the transitive closure
%    of the callee's accessible memory. All pointers to an object cease
%    to be valid when it is deallocated along with its stack frame.

%  \paragraph{Non-stack control flow}
